---
title: "Activity Monitoring"
output: html_document
---

```{r, setoptions, echo=FALSE, results=FALSE}
library(ggplot2)
library(dplyr)
library(knitr)
```

###1) Loading and preprocessing the data

I extracted the data set "activity". 
```{r}
if(!file.exists("getdata-projectfiles-UCI HAR Dataset.zip")) {
        temp <- tempfile()
        download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
        unzip(temp)
        unlink(temp)
}

activity <- read.csv("activity.csv")
```


Then, I created a new data: activity2. In this new data frame I took two steps. Firstly, I transformed the variable date as date type. Secondly, I ignored all the NA's in the original dataframe using the package "dplyr"


```{r}
activity$date<-as.Date(activity$date)
activity2<-filter(activity, steps!="NA")

```

Now, I will answer each question




### 2) What is mean total number of steps taken per day?


First, I created a dataframe of dailytotal steps
```{r}
daily<-group_by(activity2, date)
dailytotal<-summarise(daily, sum(steps))

```


Then, we can see the histogram of daily steps. 

```{r}
hist(dailytotal$`sum(steps)`, main="Daily steps", col="red", xlab = "steps", breaks = 20)
abline(v=mean(dailytotal$`sum(steps)`), col="blue", lwd=4)

```

The mean an the daily media are:

```{r}
mean(dailytotal$`sum(steps)`)
median(dailytotal$`sum(steps)`)

```





### 3) What is the average daily activity pattern?

I took the data frame activity2 and I took groups by interval. Then, I plotted the interval and the steps mean.

```{r}
intervalgroup<-group_by(activity2, interval)
intervalmean<-summarise(intervalgroup, mean(steps))
plot(intervalmean$interval, intervalmean$`mean(steps)`, type = "l")

```



### 4) Imputing missing values

With the function summary, we can see the variables and their missing values.

```{r}
summary(activity)

```

Later, I've treated the missing values assigning the mean as a replacement for NA's. Therefore, I've created a new dastaset "activity3"

"
```{r}
activity3<-activity
head(activity3)
activity3$steps[is.na(activity3$steps)]<-mean(activity$steps, na.rm = TRUE)

daily3<-group_by(activity3, date)
dailyactivity3<-summarise(daily3, sum(steps))


```

The histogram of this treated dataset is

```{r}
hist(dailyactivity3$`sum(steps)`, breaks = 20, col="red", xlab = "daily steps")
abline(v=mean(dailyactivity3$`sum(steps)`), col="blue", lwd=4)

```

And their mean and median are:
```{r}
mean(dailyactivity3$`sum(steps)`)
median(dailyactivity3$`sum(steps)`)

```




### 5) Differences in activity patterns

I've used the activity3 dataset. Then, I've created a dummy variable to identify "weekday"" and "weekend".

```{r}
activity3$day<-weekdays(activity3$date)

activity3$weekend<-ifelse((activity3$day=="sábado" | activity3$day=="domingo"),1,0)

weekend_weekday<-group_by(activity3, weekend, interval)
zz<-summarise(weekend_weekday, sum(steps))
zz1<-filter(zz, weekend==1)
zz2<-filter(zz,weekend==0)
```

Later I've created a plot that compares the steps activity in weekends and weekdays.

```{r}
par(mfrow=c(2,1), mar=c(1,1,1,1))
plot(zz1$interval, zz1$`sum(steps)`,  type = "l", main = "Weekend" )
plot(zz2$interval, zz2$`sum(steps)`, type="l", main = "weekdays")
```


